Gains and losses in rheumatoid arthritis data from inferCNV

### 2024-02-09 infercnv/reports/gains_and_losses.Rmd
delete gains_and_losses

```{r}
# function: plot gains and losses
plot_gains_and_losses <- 
  function(cnvs, annot) {
    cnvs %>%
      dplyr::rename(annotation = tidyselect::all_of(annot)) %>%
      dplyr::filter(measure == 'has', event != 'cnv') %>%
      dplyr::group_by(id, chr, event, annotation) %>%
      dplyr::summarise(n_cells = dplyr::n(),
                       n_events = sum(value),
                       prop_of_cells = n_events / n_cells) %>%
      ggplot(aes(x = id, y = chr, alpha = prop_of_cells, fill = event)) +
      geom_tile() +
      theme_classic() +
      theme(panel.spacing = unit(0, "lines")) +
      scale_fill_manual(values = c('loss' = scales::muted('blue'),
                                   'dupli' = scales::muted('red'))) +
      scale_alpha(range = c(0, 1)) +
      facet_grid(event ~ annotation)
  }

plot_gains_and_losses(cnvs, 'cell_type')
plot_gains_and_losses(cnvs, 'cluster_name')

# plot celltype
cnvs %>%
  dplyr::filter(measure == 'proportion', event == 'dupli', chr == 'chr7') %>%
  ggplot(aes(x = UMAP1, y = UMAP2, colour = cell_type)) +
  geom_point(size = 0.1) +
  theme_classic() +
  coord_fixed()

# plot subclusters on umap
cnvs %>%
  dplyr::filter(measure == 'proportion', event == 'dupli', chr == 'chr7') %>%
  ggplot(aes(x = UMAP1, y = UMAP2, colour = subcluster)) +
  geom_point(size = 0.1) +
  theme_classic() +
  coord_fixed()

# plot chr7 gain on umap
cnvs %>%
  dplyr::filter(measure == 'proportion', event == 'dupli', chr == 'chr7') %>%
  ggplot(aes(x = UMAP1, y = UMAP2, colour = value)) +
  geom_point(size = 0.1) +
  theme_classic() +
  scale_colour_viridis() +
  coord_fixed()
```

>>>>

```{r}
# function: plot gains and losses
plot_gains_and_losses <- 
  function(cnvs, annot) {
    cnvs %>%
      dplyr::rename(annotation = tidyselect::all_of(annot)) %>%
      dplyr::filter(measure == 'proportion', event != 'cnv') %>%
      dplyr::group_by(id, chr, event, annotation) %>%
      dplyr::summarise(proportion = mean(value)) %>%
      ggplot(aes(x = chr, y = id, alpha = proportion, fill = event)) +
      geom_tile() +
      theme_classic() +
      theme(panel.spacing = unit(0, "lines"),
            axis.text.y = element_text(size = 3),
            axis.text.x = element_text(angle = 270, hjust = 0, vjust = 0.5)) +
      scale_fill_manual(values = c('loss' = scales::muted('blue'),
                                   'dupli' = scales::muted('red'))) +
      scale_alpha(range = c(0, 1)) +
      facet_grid(annotation ~ event)
  }

pdf('test.pdf', height = 20)
plot_gains_and_losses(cnvs, 'cell_type')
plot_gains_and_losses(cnvs, 'cluster_name')
dev.off()

pdf('umaps.pdf', height = 15)

list(
  # plot celltype
  cnvs %>%
    dplyr::filter(measure == 'proportion', event == 'dupli', chr == 'chr7') %>%
    ggplot(aes(x = UMAP1, y = UMAP2, colour = cell_type)) +
    geom_point(size = 0.05) +
    theme_classic() +
    coord_fixed(),
  # plot chr7 gain on umap
  cnvs %>%
    dplyr::filter(measure == 'proportion', event == 'dupli', chr == 'chr7') %>%
    dplyr::arrange(value) %>%
    ggplot(aes(x = UMAP1, y = UMAP2, colour = value)) +
    geom_point(size = 0.05) +
    theme_classic() +
    scale_colour_viridis_c() +
    coord_fixed() +
    ggtitle('proportion of genes duplicated on chr7')
) %>%
  patchwork::wrap_plots(ncol = 1)

dev.off()
```
